---
title: "create-terrain-basemap"
author: Baboyma Kagniniwa | USAID
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{create-terrain-basemap}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


# Terrain Basemap for thematic mapping

Thematic maps are single-topics maps that focus on the spatial variability of a specific distribution. At USAID's Office of HIV/AIDs, we've used thematic maps to visualize the spatial distibution of HIV/AIDs Programs' key indicators at the PSNU level. ART Saturation of Nigeria at the psnu level is a good example of one of these thematic maps. In most countries, HIV/AIDs programs are implemented only in specific districts by USAID, and creating such a thematic map usually leads to having to deal with empty states. To compensate for the non-presence in some of these districts, we used terrain rasterlayer as a basemap. These basemaps are generated in way to that the hillshape is used as a beautifier of the map.

Below are how we go about creating these basemaps

```{r setup}
library(dplyr)         # General data munging and visualisation
library(ggplot2)       # General plotting
library(gisr)          # OHA/SI in house package used to create these basemap
#library(glitr)         # OHA/SI in house package used to help standardize viz
#library(sf)
```

We will be using Nigeria as our country of focus. Let's extract the country boundary and the states boundaries.

```{r}

countryname <- "Nigeria"

adm0 <- gisr::get_admin0(countries = countryname) %>% 
  dplyr::select(sovereignt, admin, iso_a2, iso_a3)

adm1 <- gisr::get_admin1(countries = countryname) %>% 
  dplyr::select(name, code_hasc)

```

What does country map look like?

```{r}
adm0 %>% gisr::gview()
```

What are the states?

```{r}
adm1 %>% 
  gisr::gview() + 
  ggplot2::geom_sf_text(data = adm1, ggplot2::aes(label = name), size = 3)
```